---
title: "Add Title"
author: "Ordinary Leading Students"
date: "6/02/2021"
header-includes:
  - \usepackage{subfig}
  - \usepackage{bbm}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r, include=FALSE, message=FALSE}

# Load libraries
library(ggplot2)        # Basic plots
library(gridExtra)      # Multiple plot same figure
library(dplyr)          # Play w/ dataframes
library(broom)          # tidy()
library(GoodmanKruskal) # nice Corr Matrix
library(Hmisc)          # compute rcorr between quantitative vars
library(corrplot)       # plot nice corr matrix
library(caret)          # trainControl (train tree predictor)
#library(performanceEstimation)

###################################################################
# SMOTE stuff
library(ROCR)

# New smote algo
#remotes::install_github("dongyuanwu/RSBID")
#library(RSBID) #cmd above to install it
###################################################################

# Set the current directory as working directory and load custom functions.
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
source("../utils/plots.R")
source("../utils/data_analysis.R")
```


\begin{abstract}
In the following paper we will present a statistical analysis of stroke related data. Our goal will be to predict whether a patient is likely to get stroke based on multiple input parameters like gender, age, various diseases, and smoking status.
**Add info**
\end{abstract}

\section{Data}

We will now load and present our dataset (that can be retrieved at the following [link](https://www.kaggle.com/fedesoriano/stroke-prediction-dataset)). \

\begin{enumerate}
\item id: unique identifier
\item gender: [Male, Female, Other]
\item age: age of the patient
\item hypertension = $\begin{cases} 1 \quad \text{if the patient has hypertension}\\  0 \quad \text{otherwise} \end{cases}$
\item heart\_disease = $\begin{cases} 1 \quad \text{if the patient has a heart disease}\\  0 \quad \text{otherwise} \end{cases}$
\item ever\_married = [Yes, No]
\item work\_type = [children, Govt\_jov, Never\_worked, Private, Self-employed]
\item Residence\_type = [Rural, Urban]
\item avg\_glucose\_level: average glucose level in blood
\item bmi: body mass index
\item smoking\_status = [formerly smoked, never smoked, smokes, Unknown\footnote{The information is unavailable for this patient}]
\item stroke = $\begin{cases} 1 \quad \text{if the patient had a stroke}\\  0 \quad \text{otherwise} \end{cases}$
\end{enumerate}\

```{r}
# Load and summarize dataset
dataset = read.csv("stroke.csv")
str(dataset)
```


\section{Data Cleaning}

We convert our variables in the right data type and we show the updated summary.\

```{r, warning=FALSE}
# Remove id column (useless)
dataset$id = NULL

# Convert qualitative variables in factors
dataset$gender          = factor(dataset$gender)
dataset$hypertension    = factor(dataset$hypertension, levels = c(0,1), labels = c("No", "Yes"))
dataset$heart_disease   = factor(dataset$heart_disease, levels = c(0,1), labels = c("No", "Yes"))
dataset$ever_married    = factor(dataset$ever_married)
dataset$work_type       = factor(dataset$work_type)
dataset$residence_type  = factor(dataset$Residence_type)
dataset$smoking_status  = factor(dataset$smoking_status)
dataset$stroke          = factor(dataset$stroke, levels = c(0,1), labels = c("No", "Yes"))

# Convert quantitative variable in numeric
dataset$bmi             = as.numeric(dataset$bmi)

# Remove duplicate column (due to uppercase)
dataset$Residence_type  = NULL 

# Show dataset summary
str(dataset)

```
\newpage

\section{ Data Visualization}

We can use different boxplots to visualize possible relations between our quantitative variables (**age**, **bmi** and **Average glucose level**) and stroke status.\

```{r, warning=FALSE, fig.width=10}
grid.arrange(ggplot(dataset, aes(x=stroke ,y=age)) +
              geom_boxplot(fill= "#FDE725FF", alpha= 0.7), 
            ggplot(dataset, aes(x=stroke, y=bmi))+
              geom_boxplot(fill= "#2D708EFF", alpha= 0.7), 
            ggplot(dataset, aes(x=stroke, y=avg_glucose_level))+
              geom_boxplot(fill= "#440154FF", alpha= 0.7), 
            ncol=3)
```

From the plots above we can infer the following:
\begin{itemize}
\item **Age** : older people are more likely to have a stroke.
\item **Bmi** :there is no evident relation between stroke and bmi.
\item **Average glucose level** : the higher the level of glucose, the higher the relation with stroke
\end{itemize}

Below we plot a few matrices to visualize possible relations between our qualitative variables and stroke status.\

```{r, warning=FALSE, fig.width=10, fig.height=7, fig.align='center'}
grid.arrange(factors_plot(tidy(table(dataset %>% select(stroke,work_type))), palette='Blues',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% select(stroke,smoking_status))), palette='Greens',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10),
            factors_plot(tidy(table(dataset %>% select(stroke,gender))), palette='Purples',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10),
            ncol=3, nrow=1)
```
```{r, warning=FALSE, fig.width=10, fig.height=5, fig.align='center'}
grid.arrange(factors_plot(tidy(table(dataset %>% select(stroke,hypertension))), palette='Greens',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% select(stroke,heart_disease))), palette='Blues',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% select(stroke,residence_type))), palette='Purples',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10),
            factors_plot(tidy(table(dataset %>% select(stroke,ever_married))), palette='Oranges',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10),
            ncol=2, nrow=2)
```

And we can finally show **qualitative** and **quantitative** correlations. \\

```{r, warning=FALSE, fig.align='center'}
qualitative_vars = c('gender', 'hypertension', 'heart_disease', 'ever_married',
                      'work_type', 'smoking_status', 'residence_type')
plot(GKtauDataframe(dataset %>% select(all_of(qualitative_vars))))
```
```{r, fig.height=6, fig.width=5, fig.align='center'}
# TODO find a way to plot table and image side-by-side

quantitative_vars = c('age', 'avg_glucose_level', 'bmi')

corr <- rcorr(as.matrix(dataset %>% select(all_of(quantitative_vars))))

corrplot(corr$r, type = "upper", tl.col = "black", tl.srt = 45)
``` 

\section{Data Manipulation}

In order to facilitate our work, we removed the level **other** from the variable gender and the level **never_worked** from the variable work_type, since they are not relevant for the analysis.\

```{r}
# Remove rows
dataset = dataset %>%
          filter(gender != 'Other') %>%
          filter(work_type != 'Never_worked')

# Remove related levels
dataset = droplevels(dataset)
```

We now have to address NA values:
```{r}
# Check for NA in all columns
for (col_name in colnames(dataset)){
   if (anyNA(dataset[[col_name]]))
    print(paste(col_name, '-> ', sum(is.na(dataset[[col_name]])), ' NA'))
}
```

We notice that we have 201 NA just in the **bmi** variable. In order to avoid losing these rows, given that we already have a small dataset, we will predict their values using a fitted tree.\

```{r, warning=FALSE, message=FALSE}

# Define test set as set with missing data
missing_index <- which(is.na(dataset$bmi))

test_set <- dataset[missing_index,]
train_set <- dataset[-c(missing_index),]

# Fit the tree
tree = caret::train(bmi ~ ., 
                    data=train_set, 
                    method="rpart", 
                    trControl = trainControl(method = "cv"))

# Replace missing data with predicted data
bmi_pred <- predict(tree, newdata = test_set)
dataset[missing_index, 'bmi'] <- bmi_pred

# clean global environment
rm(test_set, train_set, tree, bmi_pred, missing_index)
```


\section{Statistical Model}

We will now begin splitting our dataset in train/test.\

```{r}
# Set a fixed seed in order to have reproducible results
set.seed(42) # Answer to the Ultimate Question of Life, the Universe, and Everything

# What?
for (col_name in colnames(dataset)){
  levels(dataset[[col_name]]) = make.names(c(levels(dataset[[col_name]])))
}

# 50-50 split
split_train_test <- createDataPartition(y = dataset$stroke, p=0.5, list = F)
train <- dataset[split_train_test,]
test <-  dataset[-split_train_test,]

table(test$stroke)
```

\subsection{SMOTE algorithm}

Since our dataset has a problem of under sampling, we will use the SMOTE algorithm to create synthetic new data.
**ADD SMOTE EXPLANATION**\
```{r}
# New smote algo
#train_smoted <- SMOTE_NC(train, "stroke", k = 2)

# Now we have a balanced dataset
#table(train_smoted$stroke)
```

\subsection{Logistic Regression}
```{r}
Logit <- glm(stroke~., data=train, family = binomial(link = "logit"))
summary(Logit)
```
As we can see, the only significant variables are **age** and **hypertension**, 
that have both an high positive impact on the increase of the logit probability.

Also about the **smoking status** we can see, as expected, that it has a positive 
impact and a negatie one when the individual has never smoked.

\subsection{Confusion Matrix and Statistics}
```{r}
lr_prob1 <- predict(Logit, newdata = test)

# error with the for loop

#lr_preds_test <- c(0,0,0,0,0,0,0,0,0,0,0)
#i<-1
#for (thresh in seq(0.25,0.75,0.05)){
#  lr_pred <- ifelse(lr_prob1 > thresh,1,0)
#  cm <- table(
#    as.factor(lr_pred),
#    as.factor(test$stroke)
#  )[2:1, 2:1]
#  lr_preds_test[i] <- F_meas(cm) # f1 score
#  i<-i+1
#}
#names(lr_preds_test) <- seq(0.25,0.75,0.05)
#lr_preds_test
lr_pred <- as.numeric(ifelse(lr_prob1 > 0.65,"1","0"))
tb <- table(Predicted = lr_pred, Actual = test$stroke)[2:1, 2:1]
tb

#salini conf matrix
# error when knitting

#lr_pred1 <- ifelse(lr_prob1 > 0.65,"1","0")

#confusionMatrix(
#as.factor(lr_pred1),
#  as.factor(test$stroke),
#  positive = "0" 
#)
```

\subsection{ROC Curve}
```{r}
#error when knitting

#test_roc <- roc(as.numeric(test$stroke)~lr_prob1 , plot = TRUE, print.auc = TRUE,percent=TRUE, ci=TRUE)
```




