---
title: "First Paper - Stroke Analysis"
author: "Ordinary Leading Students"
date: "6/02/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes:
- \usepackage{subfig}
- \usepackage{bbm}
urlcolor: blue
---

```{r, include=FALSE, message=FALSE}

# Load libraries
library(ggplot2)        # Basic plots
library(gridExtra)      # Multiple plot same figure
library(dplyr)          # Play w/ dataframes
library(broom)          # tidy()
library(GoodmanKruskal) # nice Corr Matrix
library(Hmisc)          # compute rcorr between quantitative vars
library(corrplot)       # plot nice corr matrix
library(caret)          # trainControl (train tree predictor)
library(pROC)
#library(performanceEstimation)

###################################################################
# SMOTE stuff
library(ROCR)

remotes::install_github("dongyuanwu/RSBID")
library(RSBID) #cmd above to install it
###################################################################

# Set the current directory as working directory and load custom functions.
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
source("../utils/plots.R")
source("../utils/data_analysis.R")

```


\begin{abstract}
In the following paper we will present a statistical analysis of stroke related data. Our goal will be to predict whether a patient is likely to get stroke based on multiple input parameters like gender, age, various diseases, and smoking status.
Each row in the data provides relevant information about the patient.

After some data cleaning, we plotted some interesting variables of our dataset to get some insight on
the content and structure of our data.
We then proceeded to the analysis. 


\end{abstract}

\section{Data}

We will now load and present our dataset (that can be retrieved at the following [link](https://www.kaggle.com/fedesoriano/stroke-prediction-dataset)). \

\begin{enumerate}
\item id: unique identifier
\item gender: [Male, Female, Other]
\item age: age of the patient
\item hypertension = $\begin{cases} 1 \quad \text{if the patient has hypertension}\\  0 \quad \text{otherwise} \end{cases}$
\item heart\_disease = $\begin{cases} 1 \quad \text{if the patient has a heart disease}\\  0 \quad \text{otherwise} \end{cases}$
\item ever\_married = [Yes, No]
\item work\_type = [children, Govt\_jov, Never\_worked, Private, Self-employed]
\item Residence\_type = [Rural, Urban]
\item avg\_glucose\_level: average glucose level in blood
\item bmi: body mass index
\item smoking\_status = [formerly smoked, never smoked, smokes, Unknown\footnote{The information is unavailable for this patient}]
\item stroke = $\begin{cases} 1 \quad \text{if the patient had a stroke}\\  0 \quad \text{otherwise} \end{cases}$
\end{enumerate}

```{r}
# Load and summarize dataset
dataset = read.csv("stroke.csv")
str(dataset)
```


\section{Data Cleaning}

We convert our variables in the right data type and we show the updated summary.\

```{r, warning=FALSE}
# Remove id column (useless)
dataset$id = NULL

# Convert qualitative variables in factors
dataset$gender          = factor(dataset$gender)
dataset$hypertension    = factor(dataset$hypertension, levels = c(0,1), labels = c("No", "Yes"))
dataset$heart_disease   = factor(dataset$heart_disease, levels = c(0,1), labels = c("No", "Yes"))
dataset$ever_married    = factor(dataset$ever_married)
dataset$work_type       = factor(dataset$work_type)
dataset$residence_type  = factor(dataset$Residence_type)
dataset$smoking_status  = factor(dataset$smoking_status)
dataset$stroke          = factor(dataset$stroke, levels = c(0,1), labels = c("No", "Yes"))

# Convert quantitative variable in numeric
dataset$bmi             = as.numeric(dataset$bmi)

# Remove duplicate column (due to uppercase)
dataset$Residence_type  = NULL 

# Show dataset summary
str(dataset)

```
\newpage

\section{ Data Visualization}

We can use different boxplots to visualize possible relations between our quantitative variables (**age**, **bmi** and **Average glucose level**) and stroke status.\

```{r, warning=FALSE, fig.width=10}
grid.arrange(ggplot(dataset, aes(x=stroke ,y=age)) +
              geom_boxplot(fill= "#FDE725FF", alpha= 0.7), 
            ggplot(dataset, aes(x=stroke, y=bmi))+
              geom_boxplot(fill= "#2D708EFF", alpha= 0.7), 
            ggplot(dataset, aes(x=stroke, y=avg_glucose_level))+
              geom_boxplot(fill= "#440154FF", alpha= 0.7), 
            ncol=3)
```

From the plots above we can infer the following:
\begin{itemize}
\item \textbf{Age}: older people are more likely to have a stroke.
\item \textbf{Bmi}: there is no evident relation between stroke and bmi.
\item \textbf{Average glucose level}: the higher the level of glucose, the higher the relation with stroke
\end{itemize}

Below we plot a few matrices to visualize possible relations between our qualitative variables and stroke status.\

```{r, warning=FALSE, fig.width=10, fig.height=7, fig.align='center'}
grid.arrange(factors_plot(tidy(table(dataset %>% dplyr::select(stroke,work_type))), palette='Blues',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% dplyr::select(stroke,smoking_status))), palette='Greens',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10),
            factors_plot(tidy(table(dataset %>% dplyr::select(stroke,gender))), palette='Purples',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10),
            ncol=3, nrow=1)
```
```{r, warning=FALSE, fig.width=10, fig.height=5, fig.align='center'}
grid.arrange(factors_plot(tidy(table(dataset %>% dplyr::select(stroke,hypertension))), palette='Greens',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% dplyr::select(stroke,heart_disease))), palette='Blues',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% dplyr::select(stroke,residence_type))), palette='Purples',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10),
            factors_plot(tidy(table(dataset %>% dplyr::select(stroke,ever_married))), palette='Oranges',
                         font_count_size=3.5, font_normalized_size=5, font_percentages_size=2.5,
                         font_categories_size=10),
            ncol=2, nrow=2)
```

And we can finally show **qualitative** and **quantitative** correlations.

```{r, warning=FALSE, fig.align='center'}
qualitative_vars = c('gender', 'hypertension', 'heart_disease', 'ever_married',
                      'work_type', 'smoking_status', 'residence_type')
plot(GKtauDataframe(dataset %>% dplyr::select(all_of(qualitative_vars))))
```
```{r, fig.height=6, fig.width=5, fig.align='center'}
# TODO find a way to plot table and image side-by-side

quantitative_vars = c('age', 'avg_glucose_level', 'bmi')

corr <- rcorr(as.matrix(dataset %>% dplyr::select(all_of(quantitative_vars))))

corrplot(corr$r, type = "upper", tl.col = "black", tl.srt = 45)
``` 

\section{Data Manipulation}

In order to facilitate our work, we removed the level **other** from the variable gender and the level **never_worked** from the variable work_type, since they are not relevant for the analysis.\

```{r}
# Remove rows
dataset = dataset %>%
          filter(gender != 'Other') %>%
          filter(work_type != 'Never_worked')

# Remove related levels
dataset = droplevels(dataset)
```

We now have to address NA values:

```{r}
# Check for NA in all columns
for (col_name in colnames(dataset)){
   if (anyNA(dataset[[col_name]]))
    print(paste(col_name, '-> ', sum(is.na(dataset[[col_name]])), ' NA'))
}
```

We notice that we have 201 NA just in the **bmi** variable. In order to avoid losing these rows, given that we already have a small dataset, we will predict their values using a fitted tree.\

```{r, warning=FALSE, message=FALSE}

# Define test set as set with missing data
missing_index <- which(is.na(dataset$bmi))

test_set <- dataset[missing_index,]
train_set <- dataset[-c(missing_index),]

# Fit the tree
tree = caret::train(bmi ~ ., 
                    data=train_set, 
                    method="rpart", 
                    trControl = trainControl(method = "cv"))

# Replace missing data with predicted data
bmi_pred <- predict(tree, newdata = test_set)
dataset[missing_index, 'bmi'] <- bmi_pred

# clean global environment
rm(test_set, train_set, tree, bmi_pred, missing_index)
```


\section{Statistical Model}

We will now begin splitting our dataset in train/test.\

```{r}
# Set a fixed seed in order to have reproducible results
set.seed(36)

# 60-40 split
split_train_test <- createDataPartition(y = dataset$stroke, p=0.6, list = F)
train <- dataset[split_train_test,]
test <-  dataset[-split_train_test,]

table(test$stroke)
```

\subsection{SMOTE algorithm}

Since our dataset has a problem of under sampling, we will use the SMOTE algorithm to create synthetic new data.
**ADD SMOTE EXPLANATION**\
```{r, warning=FALSE, message=FALSE}
# New smote algo
train_smoted <- SMOTE_NC(train, "stroke", k = 2)

# Now we have a balanced dataset
table(train_smoted$stroke)
```

\subsection{Logistic Regression}

We perform a logistic regression using  \textbf{stroke} as the dependent variable and the rest of the data as independent variables in order to see if we find some significant relationships.

```{r}
Logit <- glm(stroke~., data=train, family = binomial(link = "logit"))
summary(Logit)
```
The only significant variables are **age**, **hypertension** and **average glucose level**, 
which have a positive impact on the increase of the logit probability.


\subsection{Confusion Matrix and Statistics}

```{r}
lr_prob1 <- predict(Logit, newdata = test, type = "response")

lr_preds_test <- c(0,0,0,0,0,0,0,0,0,0,0)

i<-1
for (thresh in seq(0.25,0.75,0.05)){
  lr_pred <- ifelse(lr_prob1 > thresh,1,0)
  #cm <- matrix(1:4, nrow=2)
  cm <- table(
    Predicted = factor(lr_pred, levels = c(0,1)),
    Actual = factor(test$stroke, levels = c(0,1))
  )[2:1, 2:1]
  lr_preds_test[i] <- F_meas(cm) # f1 score
  i<-i+1
}
names(lr_preds_test) <- seq(0.25,0.75,0.05)
lr_preds_test
lr_pred <- as.numeric(ifelse(lr_prob1 > 0.75,"1","0"))
tb <- table(
        Predicted = factor(lr_pred, levels = c(0,1)),
        Actual = factor(test$stroke, levels = c(0,1))
      )[2:1, 2:1]
tb
(tb[1:1,1:1] + tb[2:2, 2:2])/(tb[1:1,2:2] + tb[2:2, 1:1] + tb[1:1,1:1] + tb[2:2, 2:2]) #Accuracy
F_meas(tb) # F1
recall(tb)  # Recall
precision(tb) # Precision
```

\subsection{ROC Curve}

```{r, warning=FALSE, message=FALSE}
test_roc <- roc(as.numeric(test$stroke)~lr_prob1 , plot = TRUE, print.auc = TRUE,percent=TRUE, ci=TRUE)
```

\section{Boosting}

```{r}
for (i in 1:11) {
  levels(train_smoted[,i]) <- make.names(c(levels(train_smoted[,i])))
}
for (i in 1:11) {
  levels(test[,i]) <- make.names(c(levels(test[,i])))
}

# Tuning parameter grid
gbmGrid <- expand.grid(nIter=c(16,34,102))


trctrl <- trainControl(method = "cv"
                       , number = 3
                      # , repeats = 5
                      # , search = "random"
                       , classProbs = T
                       , summaryFunction = twoClassSummary
                       )
logit_fit <- train(stroke ~., data = train_smoted, method = "LogitBoost", trControl=trctrl,
                   tuneGrid=gbmGrid, 
                   metric= "ROC"
                   )
logit_fit
```
```{r}
plot(logit_fit)
```


