---
title: "first_lesson"
author: "Matteo Biglioli"
date: "5/24/2021"
header-includes:
  - \usepackage{subfig}
output:
  pdf_document: default
  html_document: default
---
## Data

For this work we used the dataset **Stroke Prediction Dataset** retrieved from [Kaggle](https://www.kaggle.com/fedesoriano/stroke-prediction-dataset). 

We used the dataset to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relavant information about the patient.


Below the information about the attributes:

1. id: unique identifier
2. gender: "Male", "Female" or "Other"
3. age: age of the patient
4. hypertension: 0 if the patient doesn't have hypertension, 1 if the patient has hypertension
5. heart_disease: 0 if the patient doesn't have any heart diseases, 1 if the patient has a heart disease
6. ever_married: "No" or "Yes"
7. work_type: "children", "Govt_jov", "Never_worked", "Private" or "Self-employed"
8. Residence_type: "Rural" or "Urban"
9. avg_glucose_level: average glucose level in blood
10. bmi: body mass index
11. smoking_status: "formerly smoked", "never smoked", "smokes" or "Unknown"*
12. stroke: 1 if the patient had a stroke or 0 if not

*Note: "Unknown" in smoking_status means that the information is unavailable for this patient


```{r, include=FALSE, message=FALSE}
library(ggplot2)        # Basic plots
library(gridExtra)      # Multiple plot same figure
#library(cvms)           # Confusion Matrix
library(dplyr)          # Play w/ dataframes
library(broom)          # tidy()
library(GoodmanKruskal) # nice Corr Matrix
library(Hmisc)          # compute rcorr between quantitative vars
library(corrplot)       # plot nice corr matrix
library(performanceEstimation)

###################################################################
# SMOTE stuff
library(ROCR)
###################################################################
```

<br>

## Load the dataset
Set the current directory as working directory and load the dataset.

```{r}
setwd(dirname(rstudioapi::getSourceEditorContext()$path))

source("../utils/plots.R")
source("../utils/data_analysis.R")
dataset = read.csv("stroke.csv")
str(dataset)
```

## Data Manipulation

Convert the character in factor.

```{r, warning=FALSE}
dataset$id = NULL

dataset$gender = factor(dataset$gender)
dataset$hypertension = factor(dataset$hypertension, levels = c(0,1), labels = c("No", "Yes"))
dataset$heart_disease = factor(dataset$heart_disease, levels = c(0,1), labels = c("No", "Yes"))
dataset$ever_married = factor(dataset$ever_married)
dataset$work_type = factor(dataset$work_type)
dataset$residence_type = factor(dataset$Residence_type)
dataset$Residence_type = NULL
dataset$smoking_status = factor(dataset$smoking_status)
dataset$stroke = factor(dataset$stroke, levels = c(0,1), labels = c("No", "Yes"))
dataset$bmi = as.numeric(dataset$bmi)

dataset = na.omit(dataset)
```
```{r}
str(dataset)
```

\newpage

## Data Visualization

Boxplots to visualize the stroke status in association with **age**, **bmi** and
**glucose level**.

```{r, warning=FALSE, fig.width=10}
grid.arrange(ggplot(dataset, aes(x=stroke ,y=age)) +
              geom_boxplot(fill= "#FDE725FF", alpha= 0.7), 
            ggplot(dataset, aes(x=stroke, y=bmi))+
              geom_boxplot(fill= "#2D708EFF", alpha= 0.7), 
            ggplot(dataset, aes(x=stroke, y=avg_glucose_level))+
              geom_boxplot(fill= "#440154FF", alpha= 0.7), 
            ncol=3)
```

1. **Age** there is a relation with age: older people are more likely to have a 
stroke
2. **Bmi** there is no evident relation between stroke and bmi
3. **Averagre glucose level** the higher the level of glucose, the higher the 
relation with stroke

<br>

Then we built some matrices to show the relationship between stroke and some 
**categorical variables** that have different levels.

```{r, warning=FALSE, fig.width=10, fig.height=7, fig.align='center'}
source("../utils/plots.R")
#,fig.height=12

grid.arrange(factors_plot(tidy(table(dataset %>% select(stroke,work_type))), palette='Blues',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% select(stroke,smoking_status))), palette='Greens',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10),
            factors_plot(tidy(table(dataset %>% select(stroke,gender))), palette='Purples',
                         font_count_size=4, font_normalized_size=5.1, font_percentages_size=2.5,
                         font_categories_size=10),
            ncol=3, nrow=1)
```
```{r, warning=FALSE, fig.width=10, fig.height=5, fig.align='center'}
grid.arrange(factors_plot(tidy(table(dataset %>% select(stroke,hypertension))), palette='Greens',
                         font_count_size=4.5, font_normalized_size=6, font_percentages_size=3.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% select(stroke,heart_disease))), palette='Blues',
                         font_count_size=4.5, font_normalized_size=6, font_percentages_size=3.5,
                         font_categories_size=10), 
            factors_plot(tidy(table(dataset %>% select(stroke,residence_type))), palette='Purples',
                         font_count_size=4.5, font_normalized_size=6, font_percentages_size=3.5,
                         font_categories_size=10),
            factors_plot(tidy(table(dataset %>% select(stroke,ever_married))), palette='Oranges',
                         font_count_size=4.5, font_normalized_size=6, font_percentages_size=3.5,
                         font_categories_size=10),
            ncol=2, nrow=2)
```
```{r, warning=FALSE, fig.align='center'}
# qualitative corr
qualitative_vars = c('gender', 'hypertension', 'heart_disease', 'ever_married',
                      'work_type', 'smoking_status', 'residence_type')
plot(GKtauDataframe(dataset %>% select(all_of(qualitative_vars))))
```
```{r, fig.height=6, fig.width=5, fig.align='center'}
# TODO find a way to plot table and image side-by-side

quantitative_vars = c('age', 'avg_glucose_level', 'bmi')

corr <- rcorr(as.matrix(dataset %>% select(all_of(quantitative_vars))))

corrplot(corr$r, type = "upper", tl.col = "black", tl.srt = 45)

#flattenCorrMatrix(corr$r, corr$P)
```


```{r}
# TODO clean code and stuff

# data Preprocessing, Econding with OneHotEncoding ------------------------------
# 
# dummy <- dummyVars(" ~ gender + work_type + smoking_status + ever_married + Residence_type", data=stroke)
# newdata <- data.frame(predict(dummy, newdata = stroke))
# a <- stroke[,2:4]
# b <- stroke[,8:9]
# dt <- cbind(a, b, newdata, stroke['stroke'])
# dt <- as_tibble(dt)
# 
# y <- stroke['stroke']
```
 
```{r}
# Fill missing bmi data w/ tree prediction
# TODO keep this way? should we just remove them?
# 
# missing_index <-which(is.na(dataset$bmi))
# 
# train_set <- dataset[-c(missing_index),]
# 
# tree = caret::train(bmi ~ ., 
#                     data=train_set, 
#                     method="rpart", 
#                     trControl = trainControl(method = "cv"))
# predicted_bmi <- predict(tree, newdata = dataset[missing_index,])
# 
# #######################
# # What? Why?
# # x <- mean(bmi_pred)
# # bmi_pred[202] <- x
# #######################
# 
# dataset[missing_index,"bmi"] <- predicted_bmi
# 
# dataset = na.omit(dataset)
# 
# # Check quantitative correlation is under control w/ new data
# TODO side by side :)
# 
# new_corr <- rcorr(as.matrix(dataset %>% select(all_of(quantitative_vars))))
# flattenCorrMatrix(new_corr$r, new_corr$P)
# 
# grid.arrange(corrplot(corr$r,    type = "upper", tl.col = "black", tl.srt = 45), 
#             corrplot(new_corr$r, type = "upper", tl.col = "black", tl.srt = 45), 
#             ncol=2, nrow=1)

```
```{r}
# Solve the under sampling problem with SMOTE algho to create synth new data 

# dataset <- as.data.frame(lapply(dataset, as.factor)) # What? Why? 
# 
# trainSplit <- DMwR::SMOTE(stroke ~ ., dt, perc.over = 2000, perc.under=10)
# 
# dt_synth<- rbind(trainSplit,dt)
# 
# 
# dt_synth$work_type.Never_worked <- NULL # Again, but why :D
# 
# dt_synth$avg_glucose_level <- as.numeric(dt_synth$avg_glucose_level)
# dt_synth$bmi <- as.numeric(dt_synth$bmi)
# dt_synth$age <- as.numeric(dt_synth$age)

# TODO Show some stats of new dataset
```
```{r}
 # TODO recheck correlations, maybe w/ function to plot both side-by-side
```
```{r}
 # TODO add models/predictions
```