---
title: "Second Paper"
author: "Ordinary Leading Students"
date: "6/02/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes:
- \usepackage{subfig}
- \usepackage{bbm}
urlcolor: blue
---

```{r, include=FALSE, message=FALSE}
# Loading libraries
library(ggplot2)
library(gridExtra)
library(cluster)
library(readr)
library(corrplot) 
library(factoextra)
library(summarytools)
```


\begin{abstract}
Mall dataset TO EXPLAIN

\end{abstract}

\section{Data}
We will now load and present our dataset (that can be retrieved at the following [link](https://www.kaggle.com/vjchoudhary7/customer-segmentation-tutorial-in-python)). 

```{r, message=FALSE}
setwd(dirname(rstudioapi::getSourceEditorContext()$path))

malldt<- read_csv("Mall_Customers.csv")
summary(malldt[-1])
```

\section{Data Cleaning}

```{r}
malldt$CustomerID <- NULL
malldt$Gender <- as.factor(malldt$Gender)
```

\section{Data Visualization}

```{r}
# to fix, sorry ci ho provato
grid.arrange(ggplot(malldt, aes(x=Gender, y="Annual Income (k$)")) +
              geom_boxplot(fill= "#FDE725FF", alpha= 0.7), 
             ggplot(malldt, aes(x=Gender, y="Spending Score (1-100)"))+
              geom_boxplot(fill= "#2D708EFF", alpha= 0.7), 
             ncol=2)
```


\section{Dendograms}

```{r}
# Gower distance works for mixed variables
malldt.dist<-daisy(malldt,metric="gower")

malldt.hc.com<-hclust(malldt.dist,method="complete") 
plot(malldt.hc.com, main="Complete Method") 
rect.hclust(malldt.hc.com,k=3,border=c("red","green","blue")) 

malldt.hc.sin<-hclust(malldt.dist,method="single") 
plot(malldt.hc.sin, main="Single Method") 
rect.hclust(malldt.hc.sin,k=3,border=c("red","green","blue")) 

malldt.hc.ave<-hclust(malldt.dist,method="average") 
plot(malldt.hc.ave, main="Average Method") 
rect.hclust(malldt.hc.ave,k=3,border=c("red","green","blue")) 

malldt.hc.cen<-hclust(malldt.dist,method="centroid") 
plot(malldt.hc.cen, main="Centroid Method") 
rect.hclust(malldt.hc.cen,k=3,border=c("red","green","blue")) 

malldt.hc.ward<-hclust(malldt.dist,method="ward.D2") 
plot(malldt.hc.ward, main="Ward Method") 
rect.hclust(malldt.hc.ward,k=8,border=c("red","green","blue"))  
```

\subsection{Ward Method}

The **Ward Method** is clearly the best and from it we can see that there are 8 groups.

```{r}
# Allocate obs into 8 groups
malldt.groups.ward<-cutree(malldt.hc.ward,k=8) 
malldt.groups.ward

# Number of observations in each group
table(malldt.groups.ward)
```


Below a table that shows for each group the mean value of each variable. 

```{r}
clusterdata.mean<-function(data,groups){
  aggregate(data,list(groups),function(x)mean(as.numeric(x)))
}

clusterdata.mean(malldt,malldt.groups.ward)
```


\section{K-Means algorithm}

First we have to remove the categorial variable **gender** from the code because 
the algorithm only supports numerical variables.

```{r}
malldtstd<-scale(malldt[,-1]) 
```

\subsection{Heatmap}

```{r}
#to adjust
heatmap(malldtstd)
```

\subsection{Finding K}

```{r}
set.seed(42)
k.max<-15 

wss<-sapply(1:k.max,function(k){kmeans(malldtstd,k,nstart=50,iter.max=15)$tot.withinss})

plot(1:k.max,wss,type="b",pch=19,xlab="Number of groups",ylab="Within Deviation",col="blue") 
```

Let's try with k=4 and k=6

\subsection{K = 4}

```{r}
kmeans4<-kmeans(malldt[,-1],4) 

clusplot(malldtstd, kmeans4$cluster, 
         main='2D representation of the Cluster solution',
         color=TRUE, shade=TRUE,
         labels=2, lines=0)
```

```{r}
ris4<-eclust(malldt[,-1],"kmeans",k=4) # evaluation of the clustering composition
fviz_silhouette(ris4) # dimensions and average of group's silhouette
sil4<-ris4$silinfo$widths # silhouette measure of each observation
neg_sil_index4<-which(sil4[,'sil_width']<0) # position of observation of silhouette<0
sil4[neg_sil_index4,] # observations with silhouette<0, that means that the observation should belong to the closest cluster

```


\subsection{K = 6}

```{r}
kmeans6<-kmeans(malldt[,-1],6)

clusplot(malldtstd, kmeans6$cluster, 
         main='2D representation of the Cluster solution',
         color=TRUE, shade=TRUE,
         labels=2, lines=0)

ris6<-eclust(malldt[,-1],"kmeans",k=6)
fviz_silhouette(ris6)
sil6<-ris6$silinfo$widths 
neg_sil_index6<-which(sil6[,'sil_width']<0)
sil6[neg_sil_index6,]
```














